### 2018-03-24 tex/paper.tex
completed work on background to design rationale (new sections 2.1 and 2.2), plus references; more to do

\sjtnote{References to background papers still to be added.}

\medskip
\noindent
In this section we explore the rationale for building the Marlowe language, and put it into the context of other work in the field.

\subsection{Smart contracts on blockchain}

@@

@@

Other approaches to scripting include using business process notations, such as the OMG's BPMN, process calculi such as Rholang, logic rules (in Formal Contract Logic) and finite state machines. More details of these are provided in the survey~\cite{cryptoeprint:2016:1156}.

@@

@@

Some chains have a layered design, and Cardano~\cite{cardano} does this. The settlement layer (SL) is to support settlement of transactions, whereas the computation layer (or CL) supports more powerful general computations over the chain. Moreover, the Cardano roadmap \cite{cardano-rationale} envisages the possibility of computation taking place, in part at least, on a sidechain \sjtnote{citation} rather than on the main chain itself, which may support only the SL.

\medskip
\noindent
\sjtnote{To talk about, in no particular order \ldots}
\begin{itemize}
\item
push vs pull
\item
DSLs and Embedded DSLs
\item
Multiple languages: can reference aeternity here (see presentation), and tezos (Michelson: stack-based and Liquidity, strongly typed functional)
\end{itemize}

>>>>

\sjtnote{References to some of the background work still to be added.}

\medskip
\noindent
In this section we look at existing work on smart contracts and blockchains in general, before exploring the rationale for building the Marlowe language.

\subsection{Smart contracts on blockchains}

@@

\paragraph{Multiple languages}

A common feature of many blockchain platforms is that they provide multiple scripting languages. As we noted earlier, Ethereum can be programmed at the EVM level as well as by using the high-level Solidity language. Tezos~\cite{tezos-white-paper} supports the stack-based, strongly-typed, functional language Michelson, but also provides a high-level language, Liquidity, that compiles into Michelson. Liquidity is also functional and strongly typed, but provides more constructs in a more familiar syntax, namely that of a subset of OCaml. The \ae{}ternity system~\cite{aeternity} provides multiple languages and VMs:\footnote{As presented by Erik Stenman at CODE BEAM SF, March 2018, \url{https://www.youtube.com/watch?v=VXsqvfPIdWg}} the functional language  Sophia (akin to Reason) and the Functional Typed Warded Virtual Machine (FTWVM) for safe ``system level programming'', the language Varna and the HLM for simple contracts, and a (port of) Solidity and the EVM for compatibility with Ethereum. In a similar way, Cardano provides support for IELE~\sjtnote{citation}, a rational reconstruction of the EVM, and thus for Solidity contracts too.

\paragraph{Domain-specific languages}

A domain-specific language or DSL is a high-level language designed to work in a specific field or domain. The intention is that because the users will know about the field, the constructs of the language can be designed to be meaningful to them, and also that because of its nature the DSL need not include all the features of a general purpose language. Removing this clutter and having the remaining operations directly reflect the application area is intended to make the language more accessible to domain experts who do not necessarily see themselves as programmers.

Stand-alone DSLs have the advantage of providing appropriate, domain-level error messages when things go wrong, but suffer the disadvantage of having to be implemented from scratch. An alternative to this are \emph{embedded} DSLs (EDSLs), which provide a ``little language'' for the particular domain embedded within a general-purpose host language. This means that parts of the host -- such as arithmetical expressions, or list idioms --  can be used to extend the expressibility of the DSL. A notable example of this is the financial contracts language described by Peyton Jones and his collaborators~\cite{PeytonJones:2000}, which is embedded in Haskell.

@@

There are a variety of other approaches to scripting, including business process notations, such as the OMG's BPMN; process calculi such as Rholang; logic rules (in Formal Contract Logic), and finite state machines. More details of these are provided in the survey~\cite{cryptoeprint:2016:1156}.

@@

\paragraph{UTxO and accounts}

@@

\paragraph{Interaction modality}

Contracts can be conceived of as acting in two ways. In a \emph{push} model, contracts are seen to be pro-active, and to make things happen. In the case of blockchain, this would include making payments or transactions happen. Alternatively, in the \emph{pull} model contracts enable certain things to happen, such as transactions, but the transactions need to be effected by an external player. The pull model makes fewer demands on the blockchain implementation than the push model, which requires the contract to have direct access to modify the chain.

\paragraph{Layered design and sidechains}

Some chains have a layered design, and in particular Cardano~\cite{cardano} has. The settlement layer (SL) is to support settlement of transactions but nothing more complicated, whereas the computation layer (or CL) supports more powerful general computations over the chain. Moreover, the Cardano roadmap \cite{cardano-rationale} envisages the possibility of computation taking place, in part at least, on a sidechain rather than on the main chain itself, which may support only the SL. Other off-chain approaches include the Lightning network, overlaid on blockchains (usually Bitcoin) and the channels of \ae{}ternity, introduced below.

\subsection{Designing Marlowe}

\sjtnote{To be completed.}
